---
title: "TaxCompare"
author: "Dr. Dominik Langer"
date: "Saturday, December 26, 2015"
output: html_document
---

```{r prepareStage, echo=FALSE, message=FALSE, warning=FALSE}
library(knitr) # for Knitr

library(xlsx) # for loading Excel files

library(dplyr) # for data manipulations

# For plotting
library(ggplot2) 
library(ggmap)

# For map manipulations:
library(RgoogleMaps)
library(rgdal)
library(rgeos)

dataDirectory <- "../data/"

```

## Tax scales

Tax scales for the Canton of Zurich (year: 2014) have been obtained from the canton's tax department: [data](http://www.steueramt.zh.ch/internet/finanzdirektion/ksta/de/steuerberechnung/steuertarife/_jcr_content/contentPar/downloadlist/downloaditems/steuertarife_2014_st.spooler.download.1440490483459.pdf/Steuertarife_2014_Staatssteuer_Bundessteuer.pdf)

```{r loadTaxScales, message=FALSE, warning=FALSE}

taxScaleFilePaths <- c(
      "taxScales_state_income_GT_zh_2014.csv",
      "taxScales_state_income_VT_zh_2014.csv",
      "taxScales_state_wealth_GT_zh_2014.csv",
      "taxScales_state_wealth_VT_zh_2014.csv",
      "taxScales_federal_income_GT_zh_2014.csv",
      "taxScales_federal_income_VT_zh_2014.csv"
)

for (i in seq_along(taxScaleFilePaths)) {
      if (grepl(".csv$", taxScaleFilePaths[i])) {
            tempData <- read.csv(
                  paste0(dataDirectory, taxScaleFilePaths[i]), 
                  header = FALSE,
                  stringsAsFactors = FALSE
            )
            
            fileNameComponents <- strsplit(gsub(".csv", "", taxScaleFilePaths[i]), "_")[[1]]
      
            if (fileNameComponents[2] == "state") {
                  tempData <- filter(tempData, !grepl("bis", V1)) %>%
                        select(-V1, -V4, -V6, -V7) %>%
                        rename(threshold = V2, baseTax = V3, slope = V5)
                  tempData$canton <- toupper(fileNameComponents[5])
            }
            if (fileNameComponents[2] == "federal") {
                  tempData <- select(tempData, -V4) %>% 
                        rename(threshold = V1, baseTax = V2, slope = V3)    
                  tempData$canton <- NA
            }
            
            tempData$threshold <- as.numeric(gsub("’", "", tempData$threshold))
            tempData$baseTax <- as.numeric(gsub("’", "", tempData$baseTax))
            tempData$slope <- as.numeric(gsub(".–", "", tempData$slope)) / 100
      
            tempData$law <- fileNameComponents[2]
            tempData$type <- fileNameComponents[3]
            tempData$category <- toupper(fileNameComponents[4])
            
            tempData$year <- as.numeric(fileNameComponents[6])
            
            if (i == 1)
                  taxScales <- tempData
            else
                  taxScales <- rbind(taxScales, tempData)
      }                
}

taxScaleTargetFilePath <- paste0(dataDirectory, "taxScales_cleaned.csv")
write.csv(taxScales, taxScaleTargetFilePath)
```

## Tax multipliers 

Shape file with geographical shapes of Swiss municipalities have been taken from dataset [Shape LV03](http://www.cadastre.ch/internet/kataster/de/home/services/service/plz.html).

```{r loadGISData, message=FALSE, warning=FALSE}
gisFile <- paste0(dataDirectory, "GIS_information_CH.txt")
gisData <- read.delim(gisFile, header=FALSE, fileEncoding = "UTF-8") %>%
            select(V2, V8, V3, V5, V10, V11)
names(gisData) <- c("zip", "municipality", "municipality2", "canton", "latitude", "longitude")
```

Tax multipliers for municipalities in the Canton of Zurich have been obtained from the canton's Statistical Department: [data](http://www.statistik.zh.ch/internet/justiz_inneres/statistik/de/daten/daten_oeffentliche_finanzen/gemeindesteuern/_jcr_content/contentPar/downloadlist_1/downloaditems/100_1404204330782.spooler.download.1363264483248.xls/STFUESSE.XLS)

```{r loadTaxMultipliers, message=FALSE, warning=FALSE}

multiplierFilePath <- paste0(dataDirectory, "steuerfuesse_zh.xlsx")

taxMultipliers <- read.xlsx(multiplierFilePath, 
                            sheetIndex = 4,
                            colIndex = c(4, 18),
                            header = FALSE,
                            encoding = "UTF-8",
                            startRow = 10,
                            endRow = 178
                            ) %>%
                        rename(municipality = X4, multiplier = X18) %>%
                        mutate(multiplier = multiplier / 100)
taxMultipliers$canton <- "ZH"

taxMultipliers$searchString <- sapply(strsplit(as.character(taxMultipliers$municipality), "\\."), function(x) {x[[1]][1]}) 

taxMultipliers$zip <- lapply(1:nrow(taxMultipliers), 
                             function(i) {
                                    gisSubset <- gisData[gisData$canton == taxMultipliers$canton[i], 1:3]
                                    searchString <- taxMultipliers$searchString[i]
                                    plzs <- gisSubset$zip[which(grepl(searchString, gisSubset$municipality))]
                                    if (length(plzs) == 0) {
                                          searchString <- strsplit(taxMultipliers$searchString[i], " ")[[1]][1]                                          
                                          plzs <- gisSubset$zip[which(grepl(searchString, gisSubset$municipality))]
                                    }
                                    if (length(plzs) == 0) {
                                          plzs <- gisSubset$zip[which(grepl(searchString, gisSubset$municipality2))]                                          
                                    }
                                    if (length(plzs) == 0) {
                                          print(paste("Couldn't find zip code for", taxMultipliers$municipality[i]))
                                    }
                                    plzs
                             }
                        ) 
```

## Function to calculate tax amount

```{r calculateTaxFunction, message=FALSE, warning=FALSE}

calculateTax <- function(zipCodes, amount, taxType, taxCategory, scales, multipliers) {
      selectedYear = 2014

      sapply(zipCodes, 
             function(zipCode) {
                   # Look-up canton from GIS file for filtering below
                   canton <- gisData$canton[gisData$zip == zipCode]
                   canton <- canton[!duplicated(canton)]
                   if (length(canton) > 1)
                        canton <- canton[1]
                   
                   # Scales relevant for this canton:
                   relevantScales <- filter(scales, year == selectedYear & 
                                type == taxType & 
                                canton == canton & 
                                category == taxCategory
                          )
                   baseTax <- 0
                   for(i in 1:nrow(relevantScales)) {
                        if (amount >= relevantScales$threshold[i]) {
                              baseTax <- relevantScales$baseTax[i] + 
                                    (amount - relevantScales$threshold[i]) * relevantScales$slope[i]
                        }
                   }
                   
                   multiplier <- 0
                   for(i in 1:nrow(multipliers)) {
                       if (zipCode %in% multipliers$zip[[i]]) {
                             multiplier <- multipliers$multiplier[i]
                             break
                       } 
                   } 
                   
                   (multiplier + 1) * baseTax
             }
      )
}

```

## Overlay of tax data on top of google maps

The following pages were helpful when figuring out to put overlays on GoogleMap plots: 

* [Throw some, throw some STATS on that map…(Part 1)](http://spatioanalytics.com/2013/07/12/throw-some-throw-some-stats-on-that-map-part-1/)
* [Shapefile Polygons Plotted on Google Maps Using ggmap in R – Throw some, throw some STATS on that map…(Part 2)](http://spatioanalytics.com/2014/02/20/shapefile-polygons-plotted-on-google-maps-using-ggplot-throw-some-throw-some-stats-on-that-mappart-2/)

```{r createPlot, message=FALSE, warning=FALSE}
# Let's use some test data:
income = 200000
wealth = 300000
category = "VT"

shapeDirectory <- paste0(dataDirectory, "PLZO_SHP_LV03")
shapesSwitzerland <- readOGR(shapeDirectory, layer = "PLZO_PLZ")
relevantShapes <- shapesSwitzerland[shapesSwitzerland$PLZ %in% unlist(taxMultipliers$zip),]
relevantShapes <- spTransform(relevantShapes, CRS("+proj=longlat +datum=WGS84"))
incomeTax <- calculateTax(relevantShapes$PLZ, income, "income", category, taxScales, taxMultipliers)
wealthTax <- calculateTax(relevantShapes$PLZ, wealth, "wealth", category, taxScales, taxMultipliers)
relevantShapes$tax <- incomeTax + wealthTax
relevantShapes@data$id <- rownames(relevantShapes@data)

# Convert to dataframe for plotting with ggplot - takes a while
relevantShapes.df <- fortify(relevantShapes)

# Merge tax data into shape dataframe. Hint: http://stackoverflow.com/questions/19791210/r-ggplot2-merge-with-shapefile-and-csv-data-to-fill-polygons
relevantShapes.df <- inner_join(relevantShapes.df, relevantShapes@data, by="id")


# Plot with ggplot
centerOfMap <- geocode("47.436734,8.6513793", source = "google")
googleMap <- get_map(c(lon=centerOfMap$lon, lat=centerOfMap$lat), zoom = 10, maptype = "terrain", source = "google")

ggmap(googleMap) + 
      geom_polygon(aes(x = long, y = lat, group = group, fill = tax), data = relevantShapes.df, alpha = 0.6) +
      geom_polygon(aes(x = long, y = lat, group = group), data = relevantShapes.df, colour = "black", size = 0.3, alpha = 0) +
      scale_fill_gradient2(low = "gold", mid = "grey90", high = "red", midpoint = mean(relevantShapes.df$tax))
```
